---
title: "SCRS/2021/090 Appendix B"
author: "Adrian Hordyk <adrian@bluematterscience.com>"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: yes
    toc_float: yes
    number_sections: yes
subtitle: 'North Atlantic Swordfish MSE '
always_allow_html: true
params:
  OM.n: 1
---

```{css, echo=FALSE}
body .main-container {
  max-width: 1280px !important;
  width: 1280px !important;
  }
body {
  max-width: 1280px !important;
}

```


```{r setup, include=FALSE}
library(dplyr)
library(htmltools)
library(ggplot2)
library(SWOMSE)
knitr::opts_chunk$set(echo = FALSE)

fig.num <- tab.num <- 0
fig.cap <- function() {
  fig.num <<- fig.num +1
  paste0('Figure ',  fig.num, '.')

} 
tab.cap <- function() {
  tab.num <<- tab.num +1
  paste0('Table ',  tab.num, '.')
}

OM.root <- 'G:/My Drive/1_Projects/North_Atlantic_Swordfish/OMs/grid_2021'
RepList <- readRDS(file.path(OM.root, '/OM_objects/RepList.rda'))

```

# Reference Points 

```{r}
templist <- list()
for (i in 1:length(RepList)) {
 replist <- RepList[[i]]
 derived <-  replist$derived_quants %>%
   dplyr::filter(Label %in% c('SSB_Virgin', 'SSB_MSY', 'Recr_Virgin',
                              'SPR_MSY','annF_MSY', 'Dead_Catch_MSY',
                              'Ret_Catch_MSY'))
 Kobe <- replist$Kobe %>% filter(Yr==2017)

 templist[[i]] <- data.frame(OM=i,
                             SSB0=derived$Value[1],
                             SSB_MSY=derived$Value[2],
                             R0=derived$Value[3],
                             F_MSY=derived$Value[5],
                             MSY_removals=derived$Value[6],
                             MSY_retained=derived$Value[7],
                             Dep=replist$current_depletion,
                             F_FMSY=Kobe$F.Fmsy,
                             SB_SBMSY=Kobe$B.Bmsy)
}

Ref_Points <- do.call('rbind', templist)


OM_DF$OM <- OM_DF$i
DF <- left_join(OM_DF, Ref_Points, by="OM") %>%
  select(OM, M, sigmaR, h, 'CPUE lambda' =lambda, 
         'Increase in catchability'=llq, 
         'Environmental covariate'=env, 
         SSB0, SSB_MSY, R0, F_FMSY,
         MSY=MSY_removals, 
         Depletion=Dep, F_FMSY, SB_SBMSY)

for (col in 2:7) {
  DF[,col] <- as.factor(DF[,col])
}
for (col in 8:ncol(DF)) {
  DF[,col] <- round(DF[,col],2)
}

```

`r tab.cap()` Summary table of the biological reference points for the `r nrow(DF)` operating models from the uncertainty grid. Filters at the top of each column can be used to subset the results. The table can be downloaded as a CSV file using the 'CSV' button.

```{r, results='asis'}
link <- paste0("https://github.com/ICCAT/nswo-mse/raw/master/docs/Reports/OM_Summary/Table", tab.num, ".csv")
if (knitr::is_latex_output()) {
  cat('NOTE: Dynamic tables are only available in HTML version of this report. Click [here](', link, ')^[', link, '] to download the table in CSV format (After opening link, right-click and Save As .csv).')
}
```

```{r}

ref_pointDF <- DF
par_nm <- names(ref_pointDF)[1:7]

ref_nm <- names(ref_pointDF)[8:length(names(ref_pointDF))]

col_groups <- htmltools::withTags(table(
  class = 'display',
  thead(
    tr(
      th(colspan = length(par_nm), style = "border-right: dashed 2px;", 'Parameter Values'),
      th(colspan = length(ref_nm), 'Reference Points')
    ),
    tr(
      lapply(par_nm[1:(length(par_nm)-1)], th),
      th(style = "border-right: dashed 2px;", par_nm[length(par_nm)]),
      th(shiny::HTML(paste0("SB", tags$sub("0")))),
      th(shiny::HTML(paste0("SB", tags$sub("MSY")))),
      th(shiny::HTML(paste0("R", tags$sub("0")))),
      th(shiny::HTML(paste0("F/F", tags$sub("MSY")))),
      th(shiny::HTML("MSY")),
      th(shiny::HTML(paste0("SB/SB", tags$sub("0")))),
      th(shiny::HTML(paste0("SB/SB", tags$sub("MSY"))))
    )
  )
))


DT::datatable(ref_pointDF, filter='top', 
              extensions = 'Buttons',
              options = list(
                dom = 'Blfrtip',
                buttons = list(list(extend = 'csv', filename= 'ReferencePoints')),
                pageLength = 10,
                columnDefs = list(list(width = '200px', targets = "_all"))),
              rownames = FALSE,
              escape=FALSE,
              container = col_groups) 
 
write.csv(ref_pointDF, file=file.path(getwd(), paste0('Table', tab.num, '.csv')))


```



```{r, include=FALSE}
library(GGally)

Vars <- c('M','sigmaR', 'h', 'lambda', 'llq', 'env')
Var2 <- c('Natural mortality (M)',
          'Recruitment Variability (sigmaR)',
          'Steepness (h)',
          'CPUE Lambda',
          'Increased catchability (q)',
          'Environmental covariate')

lowerfun <- function(data,mapping){
  ggplot(data = data, mapping = mapping)+
    geom_point()+
    xlim(0, Max) +
    ylim(0,Max) +
    geom_abline(intercept = 0, slope=1, linetype=2)
}

templist <- list()

for (i in seq_along(RepList)) {
  replist <- RepList[[i]]
  year <- 1950:2018
  kobe <- replist$Kobe %>% filter(Yr %in% year)
  kobe$OM <- i
  templist[[i]] <- kobe
}


kobe_DF <- do.call('rbind', templist)
kobe_DF <- kobe_DF %>% filter(Yr %in% c(2018))

OM_DF$OM <- OM_DF$i

kobe_DF <- left_join(OM_DF, kobe_DF, by='OM')

```

# SB/SB~MSY~ Correlation Plots

```{r, results='asis', message=FALSE}

VAR <- 'B.Bmsy'
LAB <- expression(SB/SB[MSY])
for (i in seq_along(Vars)) {
  
  
  nm <- Vars[i]
  name <- Var2[i]
  
  cat(paste0('\n\n## ', name, '\n\n'))
  List <- split(kobe_DF, kobe_DF[[nm]])
  tt <- list()
  store <- NA
  for (x in seq_along(List)) {
    tt[[x]] <- List[[x]] 
    ind <- match(VAR, names(tt[[x]]))
    names(tt[[x]])[ind] <- unique(List[[x]][[nm]])
    store[x] <-   names(tt[[x]])[ind]
  }
  
DF <- do.call('cbind', tt)
cols <- match(store, names(DF))
Max <- ceiling(max(DF[,cols], na.rm=T))

p1 <- ggpairs(DF[,cols],  aes(color=as.factor(DF$lambda), alpha=0.4),
              lower = list(continuous = wrap(lowerfun))) +
  xlim(0, Max) +
  theme_bw() +
  labs(x=LAB, y=LAB)
print(p1)
}

```


# F/F~MSY~ Correlation Plots

```{r, results='asis', message=FALSE}

lowerfun <- function(data,mapping){
  ggplot(data = data, mapping = mapping)+
    geom_point()+
    xlim(Min, Max) +
    ylim(Min,Max) +
    geom_abline(intercept = 0, slope=1, linetype=2)
}

VAR <- 'F.Fmsy'
LAB <- expression(F/F[MSY])
for (i in seq_along(Vars)) {
  
  
  nm <- Vars[i]
  name <- Var2[i]
  
  cat(paste0('\n\n## ', name, '\n\n'))
  List <- split(kobe_DF, kobe_DF[[nm]])
  tt <- list()
  store <- NA
  for (x in seq_along(List)) {
    tt[[x]] <- List[[x]] 
    ind <- match(VAR, names(tt[[x]]))
    names(tt[[x]])[ind] <- unique(List[[x]][[nm]])
    store[x] <-   names(tt[[x]])[ind]
  }
  
DF <- do.call('cbind', tt)
cols <- match(store, names(DF))
Min <- min(DF[,cols], na.rm=T)
Max <- max(DF[,cols], na.rm=T)

p1 <- ggpairs(DF[,cols],  aes(color=as.factor(DF$lambda), alpha=0.4),
              lower = list(continuous = wrap(lowerfun))) +
  xlim(Min, Max) +
  theme_bw() +
  labs(x=LAB, y=LAB)
print(p1)
}

```



# Timeseries Plots
For reference, the black solid line shows time-series from the 2017 stock assessment.

```{r, message=FALSE, include=FALSE}
BC_Ref <- 'G:/My Drive/1_PROJECTS/North_Atlantic_Swordfish/OMs/Grid_2020/Michael_March2020/NSWO_MSE_SS3_Base_v2/'

BCreplist <- r4ss::SS_output(BC_Ref)
Kobe <- BCreplist$Kobe %>% filter(Year %in% 1950:2018) %>% mutate(Yr=Year)
BC_TS <- BCreplist$timeseries %>% filter(Yr %in% 1950:2018)  %>%
  select(Yr, SpawnBio)
BC_TS <- BC_TS %>% mutate(Dep=SpawnBio/BCreplist$SBzero)
BC_TS <- left_join(BC_TS, Kobe, by="Yr")
BC_TS$OM <- NA

BC_TS <- BC_TS %>% tidyr::pivot_longer(cols=c(3,5,6))
BC_TS$name[BC_TS$name=='Dep'] <- 'Depletion'
BC_TS$name[BC_TS$name=='B.Bmsy'] <- 'SB/SBMSY'
BC_TS$name[BC_TS$name=='F.Fmsy'] <- 'F/FMSY'
BC_TS$name <- factor(BC_TS$name, levels=c('Depletion', 'SB/SBMSY', 'F/FMSY'),
                     ordered=TRUE)

templist <- list()
for (i in seq_along(RepList)) {
  replist <- RepList[[i]]
  Kobe <- replist$Kobe %>% filter(Yr %in% 1950:2018)
  ts <- replist$timeseries %>% filter(Yr %in% 1950:2018)  %>%
    select(Yr, SpawnBio)
  ts <- ts %>% mutate(Dep=SpawnBio/replist$SBzero)
  ts <- left_join(ts, Kobe, by="Yr")
  ts$OM <- i
  templist[[i]] <- ts
}
TS_DF <- do.call('rbind', templist)

TS_DF <- left_join(TS_DF, OM_DF, by="OM")

TS_DF <- TS_DF %>% tidyr::pivot_longer(cols=3:5)

TS_DF <- TS_DF %>% tidyr::pivot_longer(cols=c(5,6,7, ,11,12), names_to="Factor",
                                       values_to = 'Level')

TS_DF$name[TS_DF$name=='Dep'] <- 'Depletion'
TS_DF$name[TS_DF$name=='B.Bmsy'] <- 'SB/SBMSY'
TS_DF$name[TS_DF$name=='F.Fmsy'] <- 'F/FMSY'
TS_DF$name <- factor(TS_DF$name, levels=c('Depletion', 'SB/SBMSY', 'F/FMSY'),
                     ordered=TRUE)
```

```{r, results='asis', fig.width=8}

Vars <- c('M','sigmaR', 'h', 'llq', 'env')
Var2 <- c('Natural mortality (M)',
          'Recruitment Variability (sigmaR)',
          'Steepness (h)',
          'Increased catchability (q)',
          'Environmental covariate')

for (i in seq_along(Vars)) {
  
  nm <- Vars[i]
  name <- Var2[i]
  
  cat(paste0('\n\n## ', name, '\n\n'))
  
  fact <- unique(TS_DF$Factor)[i]
  j <- match(fact, Vars)
  df <- TS_DF %>% dplyr::filter(Factor==fact)

  levels <-df$Level %>% unique()
  n.level <- levels %>% length()
  temp <- list()
  for (k in 1:n.level) {
    temp[[k]] <- BC_TS
    temp[[k]]$Level <- levels[k]
  }
  BC_TS2 <- do.call('rbind', temp)

  p1 <- ggplot(df,
               aes(x=Yr, y=value, group=OM, color=as.factor(lambda))) +
    geom_line() +
    geom_line(data=BC_TS2, color='black', size=1.2) +
    expand_limits(y=c(0,1)) +
    facet_grid(name~Level,scales="free") +
    theme_bw() +
    labs(x="Year", y="Value", color='CPUE Lambda')
  print(p1)
}
```
